---
title: "Descriptive Statistics"
format: html
execute:
  echo: true
  eval: true
---


Descriptive statistics help us summarize and understand the main features of a dataset. This section covers key concepts and includes R code examples.

### ðŸ“Š Measures of Central Tendency

- **Mean**: The average of the data.
- **Median**: The middle value when the data is ordered.
- **Mode**: The most frequently occurring value.

```{r}
# Sample data
data <- c(5, 8, 9, 6, 8, 7, 10, 6, 8)

# Mean
mean(data)

# Median
median(data)

# Mode (custom function)
get_mode <- function(x) {
  uniqx <- unique(x)
  uniqx[which.max(tabulate(match(x, uniqx)))]
}
get_mode(data)
```

### ðŸ“ Measures of Dispersion

- **Range**: Difference between max and min values.
- **Variance**: The average squared deviation from the mean.
- **Standard Deviation**: Square root of the variance.

```{r}
# Range
range(data)
diff(range(data))  # Range as a single number

# Variance
var(data)

# Standard Deviation
sd(data)
```

### ðŸ“ˆ Data Visualization

Visual representations are essential for understanding the distribution of data.

```{r}
# Histogram
hist(data, main = "Histogram of Data", col = "skyblue", border = "white")
```

```{r}
# Boxplot
boxplot(data, main = "Boxplot of Data", col = "lightgreen")
```

```{r}
# Bar plot for frequency
barplot(table(data), main = "Bar Plot of Frequencies", col = "lightcoral")
```

---

### âœ… Summary

Descriptive statistics give you a first understanding of your data. You can use these techniques and R code snippets to quickly explore any dataset before moving on to more complex analysis.

More examples coming soon!
